311 Calls in 2015 
==========
*Kenneth Luna*

Going to use several packages for this project.  "plyr" allows for data summarization while ggplot improves on visualization capabilities. 

```{r}
library("forecast")
library("plyr") ## for data wrangling 
library("ggplot2") ## for data visualizations 
```

The dataset is the number of calls to 311 services in NYC pertaining to housing complaints.  The dataset is from Jan 1 '13 until today.  In aggregate, the dataset has approximately 1.4M records.  
dat
The initial dataset is a record of each call with a date and timestamp.  Preprocessing required grouping this dataset by day, resulting in about 820 records.  

```{r}
data <- read.csv("/Users/krluna/forecastingTimeseries/Project1/new_311.csv")

data$created <- as.Date(data$created, "%m/%d/%Y %H:%M:%S %p")
data <- data[-c(64),]

date <- data$created
time <- 1:length(date)
calls <- data$count
```


```{r}

par(mfrow=c(1,1))
plot(date, calls, type="l",
     xlab="Date", ylab="311 Calls", main = "Figure 1")

plot(date, calls, type="o",
     xlab="Date", ylab="311 Calls")

par(mfrow=c(1,2))
Acf(calls)
Pacf(calls)
```

Preliminary Observations 

Monthly 311 Calls from 2010 to today appears to have seasonality.  The volumne of calls to 311 appear to increase when winter arrives, while the volume drops during spring, summer and fall months.  The Monthly 311 Calls plot also shows shocks; in late 2014, there was a severe increase in the number of calls to 311.  The same holds true for the early months of 2015.  

In order to find the appropriate forecasting model, we will need to seasonally adjust this data by subtracting the monthly average across all observations.  We will take the monthly average across the entire sample there is no trend.  

```{r}

log.calls <- log(calls)

par(mfrow=c(1,1))
plot(date, log.calls, type="l",
     xlab="Date", ylab="Log(311 Calls)", main = "Figure ")

plot(date, log.calls, type="o",
     xlab="Date", ylab="311 Calls")

par(mfrow=c(1,2))
Acf(log.calls,  main = "Figure 3")
Pacf(log.calls,  main = "Figure 4")
```


Here we get monthly averages so we can seasonally adjust: 

```{r}
month <- factor(months(date), levels=format(ISOdate(2010,1:12,1),"%B"))
month.avg <- c(by(log.calls, month, mean))

print(month.avg)

avgSteps <- stepfun(1:11, month.avg, f = 0)

par(mfrow=c(1,1))
plot(avgSteps, 
     xlab = "Monthly Average", 
     ylab = "log(calls)", 
     main = "Figure 5\nMonthly Log(Calls) Average")
```

The following plot is a stepwise representation of the monthly averages across the 5+ years.  We decided to take the monthly average across all the years instead of each month-year pairing because there was trend overtime, therefore activity in a given month and year were similar to that of the same month in a different year.

Interestingly, we see spikes in calls during the winter months and some of the summer months; we assume that the calls in the winter are due to heating issues during the cold and the calls in the summer are due to noise complaints as everyone spends time outdoors.  

```{r}
log.calls.adj <- log.calls - month.avg[month]

par(mfrow=c(1,1))
plot(date, log.calls.adj, type="l",
     xlab="Date", ylab="Log(Calls) Adj.", main = "Figure 6")

par(mfrow=c(1,2))
Acf(log.calls.adj, main = "Figure 7\nLog(Calls) Adj.")
Pacf(log.calls.adj, main = "Figure 8\nLog(Calls) Adj.")
```

Based on these observations, we would suspect this is an autoregressive model as the correlogram gradually dies down while the partial coorelogram cuts off after lag-1.  

```{r}
diff.log.calls.adj <- c(NA, diff(log.calls.adj))
diff2.log.calls.adj <- c(NA, diff(diff.log.calls.adj))
```

Differencing: 

```{r}
par(mfrow=c(1,1))
plot(date, diff.log.calls.adj, type="l",
     xlab="Date", ylab="Diff. Log(Calls) Adj.")

par(mfrow=c(1,2))
Acf(diff.log.calls.adj, main = "Figure 9\nDifference-1")
Pacf(diff.log.calls.adj, main = "Figure 10\nDifference-1")
```

```{r}
par(mfrow=c(1,1))
plot(date, diff2.log.calls.adj, type="l",
     xlab="Date", ylab="Diff2. Log(Calls) Adj.")

par(mfrow=c(1,2))
Acf(diff2.log.calls.adj, main = "Figure 11\nDifference-2")
Pacf(diff2.log.calls.adj,  main = "Figure 12\nDifference-2")
```

Based on our observations of the 1st and 2nd difference of the seasonally adjust log(Calls), it appears that there is no need to difference.  Difference-1 and Difference-2 both have lag-1 values that are fairly negative, indicating over-differencing.  

```{r}

fit.00 <- Arima(log.calls.adj, c(0, 0, 0), include.constant=FALSE)
fit.01 <- Arima(log.calls.adj, c(0, 0, 1), include.constant=FALSE)
fit.02 <- Arima(log.calls.adj, c(0, 0, 2), include.constant=FALSE)
fit.10 <- Arima(log.calls.adj, c(1, 0, 0), include.constant=FALSE)
fit.11 <- Arima(log.calls.adj, c(1, 0, 1), include.constant=FALSE)
fit.12 <- Arima(log.calls.adj, c(1, 0, 2), include.constant=FALSE)
fit.20 <- Arima(log.calls.adj, c(2, 0, 0), include.constant=FALSE)
fit.21 <- Arima(log.calls.adj, c(2, 0, 1), include.constant=FALSE)
fit.22 <- Arima(log.calls.adj, c(2, 0, 2), include.constant=FALSE)

# With constant:
fit.00c <- Arima(log.calls.adj, c(0, 0, 0), include.constant=TRUE)
fit.01c <- Arima(log.calls.adj, c(0, 0, 1), include.constant=TRUE)
fit.02c <- Arima(log.calls.adj, c(0, 0, 2), include.constant=TRUE)
fit.10c <- Arima(log.calls.adj, c(1, 0, 0), include.constant=TRUE)
fit.11c <- Arima(log.calls.adj, c(1, 0, 1), include.constant=TRUE)
fit.12c <- Arima(log.calls.adj, c(1, 0, 2), include.constant=TRUE)
fit.20c <- Arima(log.calls.adj, c(2, 0, 0), include.constant=TRUE)
fit.21c <- Arima(log.calls.adj, c(2, 0, 1), include.constant=TRUE)
fit.22c <- Arima(log.calls.adj, c(2, 0, 2), include.constant=TRUE)

models <- data.frame(p = rep(c(0, 0, 0, 1, 1, 1, 2, 2, 2), 2),
                     d = rep(0, 18),
                     q = rep(c(0, 1, 2), 6),
                     include.constant = c(rep(FALSE, 9), rep(TRUE, 9)),
                     loglik = c(fit.00$loglik, fit.01$loglik, fit.02$loglik,
                                fit.10$loglik, fit.11$loglik, fit.12$loglik,
                                fit.20$loglik, fit.21$loglik, fit.22$loglik,
                                fit.00c$loglik, fit.01c$loglik, fit.02c$loglik,
                                fit.10c$loglik, fit.11c$loglik, fit.12c$loglik,
                                fit.20c$loglik, fit.21c$loglik, fit.22c$loglik),
                     aicc = c(fit.00$aicc, fit.01$aicc, fit.02$aicc,
                                fit.10$aicc, fit.11$aicc, fit.12$aicc,
                                fit.20$aicc, fit.21$aicc, fit.22$aicc,
                                fit.00c$aicc, fit.01c$aicc, fit.02c$aicc,
                                fit.10c$aicc, fit.11c$aicc, fit.12c$aicc,
                                fit.20c$aicc, fit.21c$aicc, fit.22c$aicc)
                     )
```

The p, d, q selected based on AICC is the ARIMA(2, 0, 2) without a constang with an AICC of -166.120. 

Here is the estimated model: 
```{r}

fit <- Arima(log.calls.adj, c(2,0,2), include.constant = FALSE)
print(fit)

(1-pnorm(abs(fit$coef)/sqrt(diag(fit$var.coef))))*2

```

x(t) = .4622 * x(t-1) + .3911 * x(t-2) + eps(t) + .2932 * eps(t-1) + -.5413 * eps(t-2)

```{r}

resid <- residuals(fit)

lb12 <- Box.test(resid, lag=12, type = "Ljung-Box", fitdf=4)
lb24 <- Box.test(resid, lag=24, type = "Ljung-Box", fitdf=4)
lb36 <- Box.test(resid, lag=36, type = "Ljung-Box", fitdf=4)
lb48 <- Box.test(resid, lag=48, type = "Ljung-Box", fitdf=4)

```


P-Values are not statistically significant, therefore indicating that we have removed any structure in the data and that it is close to white noise.  

```{r}
par(mfrow=c(1,1))
plot(date, resid, type="l",
     main = "Figure 13")

abline(h=0.003537943, col = 'red', lty=2)

par(mfrow=c(1,2))
Acf(resid, main = "Figure 14\nResiduals")
Pacf(resid, main = "Figure 15\nResiduals")
```


```{r}

forecast(fit, h = 50, level = 95)

```


```{r}
par(mfrow=c(1,1))
plot(forecast(fit, h=50, level = 95),
     main = "Figure 16\nForecasts from Arima(2,0,2) with zero mean")
```


