Homework 9
==========
*Kenneth Luna*

For this assignment, we will use the `forecast` library.
```{r}
library("forecast")
```

We will analyze the `gilt` dataset:
```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/gilt.csv")

long <- data$long # 20 year UK gilt interest rate
short <- data$short # 91 day UK Treasury bill interest rate

n <- nrow(data)
time <- 1:n
date <- as.Date(data$date)
```


Problem 1
---------

Here is the UK interest rate spread:

```{r}
spread <- long - short
par(mfrow=c(1,1))
plot(date, spread, type="l",
     xlab="Date", ylab="Long-Short Spread")
```

**Does the series appear to be stationary?**

This plot does not appear to be stationary. Although the mean remains constant, the variance seems to vary significantly.  

Here are an ACF and PACF for the spread:

```{r}
par(mfrow=c(1,2))
Acf(spread)
Pacf(spread)
```

**Based on the ACF and the PACF of the raw data, explain why an AR(1) model
might be reasonable.**

have to find reason for this **** 

Problem 2
---------

Here we estimate an AR(1) model via least squares regression:

```{r}
# Uncomment the following command to create a lagged-version of the
# spread variable, then use the "lm" command to perform a least
# squares regression fit.

lag.spread <- c(NA, spread)[1:n]
fit.lm <- lm(spread ~ lag.spread)   #???? # replace with appropriate call to "lm" command

# Use the "summary" command to see the standard errors from the regression
# fit.
summary(fit.lm) # Uncomment this line
```


Here is the t-statistic for testing H0: rho = 1 against H1: rho < 1:

```{r}
(tau.mu.spread <- (0.85771 - 1) / (0.04512))
# add code to compute the t statistic.  You will need to copy and paste
# the appropriate values from the "summary(fit.lm)" output.
```

Here is the p-value for the test:
```{r}
# Use the t-statistic to compute the pvalue.  You can either use a z table
# to compute an approximate p-value by hand, or you can use the "pt" function
# to compute an exact p-value, specifying "df=n-2".  (Make sure you compute
# the correct tail area; "pt(tstat, df)" gives the probability that a t random
# variable is less than the value "tstat".
print(n)
```

**Does this p-value provide strong evidence against the random walk
hypothesis?**

Since tau.mu is not great than -2.57 and less than -.42, we reject the null hypothesis in this case. 

**Is rho-hat significantly less than 1 at level .01?**



Problem 3
---------

Here, we look at the differenced data:

```{r}
# construct a "diff.spread" variable containing the spread differences
diff.spread <- diff(spread)
```

Here are the ACF and PACF:

```{r}
par(mfrow=c(1,2))
Acf(diff.spread)
Pacf(diff.spread)
```

Here are the sample mean and sample sd
```{r}
mean(diff.spread)
sd(diff.spread)
# Compute and print the sample mean and sd of "diff.spread" variable.  You
# may need to specify "na.rm=TRUE" when you use these functions.
```

**Based on the ACF, PACF, sample mean and sample standard deviation of the
differenced data, argue that a random walk without drift might also provide a
reasonable model for the UK interest rate spread.**



Problem 4
---------

**Perform the Dickey-Fuller test (tau[mu]) of driftless random walk versis
stationary AR(1).**


**Compute the approximate p-value for the AR(1) hypothesis, based on the
Dickey-Fuller table.**



**Is rho hat significantly less than 1 at level .01?**



**Accordingly, based on a hypothesis test at level .01, decide whether the UK
interest rate spreads are driftless random walk or stationary AR(1).**


**This conclusion is different from the one you arrived at in Problem 1.  What
is the reason for the difference?**


**Which conclusion is more justifiable, on statistical grounds?**



**If you used a significance level of .05, what would you conclude from the
Dickey-Fuller test?**

