Homework 2: Problems 1--3
=========================
Kenneth Luna


---


In Problems 1--3, we analyze the Russell 2000 Stock Index, adjusted daily
closing price, recorded from 10 September 1987 to 11 February 2015.  

```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/russell.csv")
russell <- data$russell
date <- as.Date(data$date)
time <- 1:length(russell)
```

We will work with the mean-adjusted series:

```{r}
today <- russell - mean(russell)
```


Problem 1
=========

Part A
------

Here is a plot of Today's Russell versus time, as well as Yesterday's Russell
versus time:

```{r}
yesterday <- c(NA, today[-length(today)])

library("ggplot2")

ggplot() + geom_line(aes(date, today, linetype = "dotted")) + geom_line(aes(date, yesterday, linetype = "dashed")) + 
  scale_linetype_manual(name = "Legend", values=c("dotted", "dashed"), labels = c("Today", "Yesterday")) + theme(legend.position = "bottom") + 
  ylab("Mean-Adjusted Russell") + xlab("Date")
```

Next, here is a plot of Today's Russell versus time, as well as
0.5 (Yesterday's Russell):


```{r}
ggplot() + geom_line(aes(date, today, linetype = "dotted")) + geom_line(aes(date, 0.5 *yesterday, linetype = "dashed"))  + 
  scale_linetype_manual(name = "Legend", values=c("dotted", "dashed"), labels = c("Today", "(.05) * Yesterday")) + theme(legend.position = "bottom") + 
  ylab("Mean-Adjusted Russell") + xlab("Date")
```


Part B
------

**Based on these two plots, which seems to be a better forecast of Today's
Russell: Yesterday's Russell, or 0.5 (Yesterday's Russell)?**

Yesterday's Russell is a better forecast of Today's Russell. Where the Russell Index is today is a strong indicator of where it will likely be tomorrow.  Therefore, adjusting for Yesterday's Russell is highly correlated to Today's Russell.  0.5 (Yesterday's Russell) under-estimates both the drops and increases in the actually RI.  


Part C
------

Here are the average squared forecast errors for the two forecasts:

```{r}

Error1 <- mean((yesterday-today) ^ 2, na.rm = TRUE)
Error2 <- mean(((0.5 * yesterday) - today) ^ 2, na.rm = TRUE)

Error1
Error2
# Replace this comment with code to compute average squared forecast
# errors for the two forecasts (you will compute two numbers: one for
# 'yesterday', and one for '0.5 yesterday').
#
# Hint: use the mean function.
```

**Based on this, which one was better?**

Yesterday's Russell is better at forecasting Today's Russell.  The average squared error is significantly lower versus that of (0.5 * Yesterday's Russell).  


Problem 2
=========

Part A
------

Here is a plot of Today's Russell versus Yesterday's Russell:

```{r}
ggplot() + geom_point(aes(yesterday, today), alpha = .5) + ylab("Today") + xlab("Yesterday")
```

**Describe any patterns you see.**

There appears to be a very strong positively linear relationship.  If we were to draw a straight line to explain the relationship, we would realize some noise (variance between expected Y versus actual Y.  Fortunately, the variance is limited in size.  Thus I will conclude that Yesterday's Russell Index is a very good indicator of Today's Russell Index.  

Part B
------

Here is a linear regression fit of Today's Russell on Yesterday's Russell:

```{r}
model <- lm(today ~ yesterday)
summary(model)
```

**What is the prediction of Today's Russell implied by the regression
coefficients?**

The prediction based on the regression coefficients is the following: 

$$today = .9999972 * yesterday + .1493892$$

**Is this consistent with your answers to Problem 1, parts B and C?**

Yes this results holds consistent with my findings in Problem 1.  Yesterday's Russell Index is highly correlated with Today's Russell Index.  




Part C
------

The 95% confidence interval for the slope is
```{r}
confint(model, "yesterday", level = .95)
```

**Is the slope in your fitted regression significantly different from 1?**

No, the slope in my fitted regression is not significantly different from 1.  It is very close to 1.  Again, this is due to the fact that yesterday's Russell Index is a strong indicator of  Today's Russell Index.


The 95% confidence interval for the intercept is
```{r}
confint(model, "(Intercept)")
```

**Briefly comment on the intercept as well.**

This intercept value has limited reliability based on high p-value.  Based on its 95% confidence interval, it appears that the intercept is attempting to account for any noise between yesterday's RI versus today's RI.  


Part D
------
**Based on everything you have done so far, do you see any strong evidence
that the Russell is *not* a random walk**?

Based on this analysis, there is a strong amount of evidence that the Russell Index is not a random walk.  Random Walk assumes that changes in the RI have the same distribution or likelihood but based on these results, it is apparent that is not true.  What occurs in the RI on day t is strong reflected in what occurred on day day t-1.  

Part E
------
Here is the correlation coefficient between Today's Russell and Yesterday's Russell:

```{r}
cor(today, yesterday, use = "complete.obs")
```

**Based on this, how strong is the linear association between Today's Russell
and Yesterday's Russell?**

The linear relationship is very strong as the absolute value of the correlation coefficient is close to 1.  

Problem 3
=========

Part A
------

We define the Russell returns as
```{r}
ret.today <- diff(russell) / russell[2:length(russell)]
ret.yesterday <- c(NA, ret.today[-length(ret.today)])
```


Here is a plot of Today's return versus time:
```{r}
newdate <- date[2:length(date)]
ggplot() + geom_point(aes(newdate, ret.today), alpha = .5) + xlab("RI Return") + ylab("Date")
```


Here are the sample average and standard deviation of the returns:
```{r}
mean(ret.today)
sd(ret.today)
```


Here is an ordinary t-test of whether the mean returns are zero:
```{r}
t.test(ret.today)
```

**Are the mean returns significantly different from zero?  Interpret your
findings.**

Both mean results are significantly close to 0.  Based on these results, we can be 95% percent confident that the mean return on the RI is 0 in our  dataset.  

Part B
------

Here is a histogram of the Russell returns:
```{r}
ggplot() + geom_histogram(aes(ret.today), stat = "bin", binwidth = .005) + 
  ylab("Count") + xlab("RI Return")

```

Here is a boxplot:
```{r}
ggplot() + geom_boxplot(aes(x = factor(0), y = ret.today)) + coord_flip() + xlab("") + scale_x_discrete(breaks = NULL) + ylab("RI Returns")
```

Finally, here is a normal quantile-quantile plot:
```{r}
qplot(sample = ret.today, stat="qq")
```

**Do you think that the Russell returns are normally distributed?  Explain.**

The Russell returns appear to be normally distributed.  First, the histogram is bell shaped, with a high concentration of results tethered to 0, the mean.  The boxplot helps visualize some of the outliers.  It appears that RI Return values on the negative end were a bit more severe versus those that were positive.  Still, the 25% and 75% quartiles are very similar in range.  The Q-Q plot also appears very symmetrical in its shape and its fat tails.  



Part C
------

Here is a plot of today's return versus yesterday's return
```{r}
ggplot() + geom_point(aes(ret.today, ret.yesterday), alpha = .5, size = 1) + ylab("Yesterday's Return") + xlab("Today's Returns")
```

Here is a zoomed in plot:
```{r}
ggplot() + geom_point(aes(ret.today, ret.yesterday), alpha = .5, size = 1) + ylab("Yesterday's Return") + xlab("Today's Returns") + 
  xlim(-.05,.05) + ylim(-.05,.05)
```


**Does this plot appear very different from the one in Problem 2(A)?  Which
seems to be easier to predict: Today's Russell, or Today's return?**

The plot appears very different when compared to Problem 2(A).  Because the graph in 2(A) looked at absolute return, there was a linear trend when comparing yesterday versus today.  In this graph, we are looking at relative return; because the value of the Russell Index today is very close to the value yesterday, we can infer that the return will be 0.  Both visualizations point to the same conclusions.  

Part D
------

Here is a linear regression of today's returns on yesterday's returns:
```{r}
model <- lm(ret.today ~ ret.yesterday)
summary(model)
```

**What is the prediction of today's return implied by the regression
coefficients?**

The prediction of today's return implied by the regression coefficients is the following: 

$$ret.today = ret.yesterday * 0.0021011 + .0001942$$

**Are the coefficients statistically significantly different from zero?**

The coefficients are not statistically significantly different from zero because the p-values are greater than the .05 threshold. 



