
##### Kenneth Luna
##### Homework1


In this problem we will consider the daily U.S. Dollar-Euro exchange rate,
daily, 4 Jan 1999 to 27 Jan 2015 (*n* = 4115).

```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/euro.csv")
date <- as.Date(data$date)
euro <- data$euro
```


Part A
------

Here is a time-series plot of Euro.

```{r}
library(ggplot2)
ggplot() + geom_line(aes(date, euro)) + xlab(NULL) + ylab("Rate") + labs(title = "U.S. Dollar-Euro Exchange Rate")
```


**Does a straight-line model seem appropriate?**
A straight-line model does not seem appropriate here because there is too much volatility in the US Dollar-Euro exchange rate.  Furthermore, there isn't a constant long term trend or even a seasonal component.  



Part B
------

Here, we fit two models and predict the value at 28 Jan 2015 (time = 4116).  In
the first model, we fit to observations 1 to 700.  In the second model, we fit to observations 701 to 4115.

```{r}
time <- 1:length(euro) # create the time variable

model1 <- lm(euro ~ time, subset=1:700)
summary(model1)

model2 <- lm(euro ~ time, subset=701:4115)
summary(model2)
```

Here are the predicted values and 95% prediction intervals for the two models:
```{r}
newdata <- data.frame(time = 4116)
predict(model1, newdata, interval = 'prediction', level = .95)
predict(model2, newdata, interval = 'prediction', level = .95)
```

**Did both of the forecast intervals succeed in containing the actual value
for 28 Jan 2015?  If not, then use what you learned in Problem~1 to give a
statistical explanation of what went wrong.**

The USD/Euro exchange rate on Jan 28 was 1.134 at close of business day.  Neither model succeeded in containing a 95% confidence interval that captured that end of day exchange rate.  Model1 was significantly off the mark; this is likely due to a lack of and less relevant time series data as it mostly pertains to the early 2000s.  Model2 was closer but over-estimated the exchange rate.  

The issue with this model is that it is not linear.  Unlike the first problem where GDP clearly grew over time, the exchange rate of USD to Euros shows different characteristics in terms of its trends.  A linear regression model fails to capture this, and thus fails to accurately predict these exchange rates.  

Part C
------

Here, we superimpose the fitted lines from both models on the time series plot.

```{r}
cols <- c("Actual" = "black", "Fit to 1-700" = "red", "Fit to 701-4115" = "blue")

ggplot() + geom_line(aes(time, euro, colour = 'Actual'), alpha = .75) + 
  
  geom_abline(aes(colour = "Fit to 1-700"), intercept = 1.104e+00, 
              slope = -3.873e-04, size = 1, linetype = 'dashed') + 
  
  geom_abline(aes(colour = "Fit to 701-4115"), intercept = 1.054e+00, 
              slope = 9.265e-05,  size = 1, linetype = 'dashed') +
  scale_colour_manual(name = "Legend", values= cols) + 
  xlab(NULL) + ylab("Rate") + labs(title = 'Actual Data with Models 1 & 2') + 
  theme(legend.position = "bottom")
```
