

##### Kenneth Luna
##### Homework1

In this problem we will investigate the United States Gross Domestic Product 
(GDP).  The data file `gdp.csv` contains the real US GDP, quarterly, from 1946, 
Q4 to 2014 Q2, seasonally adjusted and inflation adjusted, in billions of 2009 
dollars.  



Part A
------

We read the data file and extract the `date` and `gdp` columns with 
the following commands:

```{r}
library(ggplot2)
data <- read.csv("http://ptrckprry.com/course/forecasting/data/gdp.csv")
date <- as.Date(data$date)
gdp <- data$gdp
```

Here is a plot of the GDP series versus date:

```{r}
ggplot() + geom_line(aes(date, gdp)) + ylab("GDP") + xlab(NULL) + labs(title = "US GDP from Q1'46 to Q2'14")
```

**Does GDP seem to grow linearly over time?**

Yes, it appears GDP grows linearly over time based on what this graph looks like. 



Part B
------

Here is a plot of log(GDP).

```{r}
log.gdp <- log(gdp) # create the log(GDP) variable
ggplot() + geom_line(aes(date, log.gdp)) + ylab("Log(GDP)") + xlab(NULL) + labs(title = "Log(US GDP) from Q1'46 to Q2'14")
                                              
```

**Does log GDP appear to grow linearly over time?  If so, then what does this
imply about the growth of GDP itself?**

Log GDP grows linearly over time.  What we can infer from this model is that the log straightens out any exponential growth and reduces variance from our prior time series model.  Considering that this model appears significantly more linear compared to GDP itself, we can imply that growth is actually exponential.  


Part C
------

We fit an ordinary linear regression model for log(GDP), using time as the 
predictor variable.

```{r}
log.gdp <- log(gdp)

time <- 1:length(gdp) # create the time variable
model <- lm(log.gdp ~ time)
model
```

Based on the fitted model, we forecast the log(GDP) for the third quarter
of the year 2014 (time = 272).  

```{r}
newdata <- data.frame(time = 272)
predict(model, newdata)
```

We also construct a 95% prediction interval to go with the
point forecast:

```{r}
predict(model, newdata, interval = 'prediction', level = .95)
```

**Do you think this interval forecast (prediction interval) is valid?  Does it
seem too wide, or too precise?  Explain.**

I think this interval forecast is too wide.  If you convert to the real GDP, you get a 95% confidence interval range of 16,809 to 21,262, which is too severe based on historical trends.  For the GDP to drop down to those lower levels, we must assume there is a severe economic shock around the corner.  Although that might be true, what is the likelihood that it is?  The same can be said about the high end of the CI; the US economy has been recovering for the last 8 years, and it would assume that our best year is still to come which is highly doubtful.  I think we could generate a more precise confidence interval.  

Part D
------

Next is a plot of the log(GDP) series with the fitted line superimposed.

```{r}
ggplot() + geom_line(aes(date, log.gdp)) + geom_abline(intercept = 8.400e+00 , slope = 8.857e-05, color = 'red', linetype = 'dashed') + ylab("Log(GDP)") + xlab(NULL) + labs(title = "Log(US GDP) with Model")
```

**Does the line fit well?**

The line fits well up until the last 6 to 7 years.  The same can be said about the front end of the curve around 1950.  The fitted line seems to over-state GDP relative to actual GDP. 

Part E
------

Here is a plot the residuals from the fitted line versus time.

```{r}
res <- residuals(model)
ggplot() + geom_point(aes(time, res), colour = 'red') + labs(title = "Residuals Over Time") + ylab("Residuals") + xlab(NULL)
```

**What potential problems with the linear model are indicated by this plot?
Do you think these problems could spoil the validity of the forecast
interval?**

The problem with this model is that each observation is not entirely indepenent of one another.  Based on the residual plot, there is a very predictable pattern; in years when GDP grew, it is highly probably that the subsequent year will realize an increase as well and vice-versa.  

This observation can spoil the validity of our forecast model as it violates a key assumption of linear regression that each observation is independent.  









