Project 2
==========
*Kenneth Luna*

```{r}
library("forecast")
library("tseries")
```

```{r}
data <- read.csv("tsla.csv")
data$Date <- as.Date(data$Date, "%d-%b-%y")
data <- data[order(data$Date),]

date <- as.Date(data$Date, "%d-%b-%y")
time <- 1:length(date)
tsla <- data$Close
```


```{r}
par(mfrow=c(1,1))
plot(date, tsla, type="l",
     xlab="Date", ylab="Closing Price", main = "Figure 1\nTSLA Price")
```


Convert closing stock price to log.tsla.

```{r}
log.tsla <- log(tsla)
```


Problem 1
---------

Here is a time series plot of log.tsla, along with the ACF and PACF:

```{r}
par(mfrow=c(1,1))
plot(date, log.tsla, type="l",
     xlab="Date", ylab="log.tsla", main = "Figure 2\nLog(TSLA)")

par(mfrow=c(1,2))
Acf(log.tsla, main = "Figure 3\nLog(TSLA)")
Pacf(log.tsla, main = "Figure 4\nLog(TSLA)")
```

Here is the first difference, along with the ACF and PACF:

```{r}
## Add code to compute the first difference and make the plots.
diff.log.tsla <- c(NA, diff(log.tsla))

par(mfrow=c(1,1))
plot(date, diff.log.tsla, type="l",
     xlab="Date", ylab="1st-Difference Log TSLA", main = "Figure 5\n1st-Diff. Log(TSLA)")

par(mfrow=c(1,2))
Acf(diff.log.tsla, main = "Figure 6\n1st-Diff. Log(TSLA)")
Pacf(diff.log.tsla, main = "Figure 7\n1st-Diff. Log(TSLA)")
```

Here is the second difference, along with the ACF and PACF:

```{r}
## Add code to compute the second difference and make the plots.
diff2.log.tsla <- c(NA, diff(diff.log.tsla))

par(mfrow=c(1,1))
plot(date, diff2.log.tsla, type="l",
     xlab="Date", ylab="2nd-Difference Log TSLA", main="Figure 8\n2nd-Diff. Log(TSLA)")

par(mfrow=c(1,2))
Acf(diff2.log.tsla, main = "Figure 9\n2nd-Diff. Log(TSLA)")
Pacf(diff2.log.tsla, main = "Figure 10\n2nd-Diff. Log(TSLA)")
```

**Does the series appear to be stationary?**


**Can you identify an ARIMA(p,d,q) model from these plots?**

Similar to the rupee data, an ARIMA(0,1,0) seems to be the best fit as it shows now signs of autocorrelation.  


Problem 2
---------

Here are the AICc for all ARIMA(p,1,q) models without constant for p and q
ranging from 0 to 2:

```{r}
d <- 1
aiccMin <- 0
# choose p, q with AICc
for (include.constant in c(FALSE, TRUE)) {
    for (p in 0:4) {
        for (q in 0:4) {
            # work-around bug in R by manually differencing
            fit <- Arima(diff(log.tsla), c(p,0,q), include.constant=include.constant, method="ML")
            cat("ARIMA(", p, ",", d, ",", q, ")(constant=", include.constant, ") : ", fit$aicc, "\n", sep="");
    
            if (fit$aicc < aiccMin)
              aiccMin <- fit$aicc      
        }
    }
}


print(aiccMin)
```


**Select an ARIMA(p,1,q) model.**

The best fitting model is the ARIMA(0,1,0) based on the AICC value of -4676.754.  This model does consist of a constant.  This holds consistent with our preliminary observation.  

Here is code to fit the model, then compute residuals and the fitted values:
```{r}
(fit_mean <- Arima(log.tsla, c(0,1,0), include.constant=TRUE, method="ML"))
```

Here are the residuals, with the last 10 residuals printed out:
```{r}
resid <- as.numeric(residuals(fit.mean))
tail(resid, n=10)
```

Here are the fitted values, with the last 10 fitted values printed out:
```{r}
f <- fitted.values(fit.mean)
tail(f, n=10)
```

Here is the one step ahead forecast and 95% forecast interval:

```{r}
forecast(fit.mean, h = 1, level = 95)
```

Problem 3
---------

Here is a plot of the residuals:
```{r}
# Add plot of the residuals.
par(mfrow=c(1,1))
plot(time, resid, type="l",  main = "Figure 11\nResiduals")
```

Here are an ACF and PACF of the residuals:

```{r}
par(mfrow=c(1,2))
Acf(resid,  main = "Figure 12\nACF Residuals")
Pacf(resid,  main = "Figure 13\nPACF Residuals")
```

Here are an ACF and PACF of the squared residuals:

```{r}
par(mfrow=c(1,2))
Acf(resid^2, main = "Figure 14\nACF Residuals^2")
Pacf(resid^2, main = "Figure 15\nACF Residuals^2")
```


Problem 4
---------

Here are the AICc values for the ARCH(q):

```{r}
q <- 0:10
loglik <- rep(NA, length(q))
N <- length(resid)

for (i in 1:length(q)) {
    if (q[i] == 0) {
        loglik[i] <- -0.5 * N * (1 + log(2 * pi * mean(resid^2)))
    } else {
        fit <- garch(resid, c(0,q[i]), trace=FALSE)
        loglik[i] <- logLik(fit)
    }
}

k <- q + 1
aicc <- -2 * loglik  + 2 * k * N / (N - k - 1)

print(data.frame(q, loglik, aicc))
```

Here is the AICc for the GARCH(1,1):

```{r}
fit <- garch(resid, c(1,1), trace=FALSE)
loglik <- logLik(fit)
k <- 2
aicc <- -2 * loglik  + 2 * k * N / (N - k - 1)

print(data.frame(loglik, aicc))
```


Here are the summary and log likelihood of the selected model:

```{r}
fit.var <- garch(resid, c(0,5), trace=FALSE)
summary(fit.var)
logLik(fit.var)
```


**Comment on the statistical significance of the parameter values of your selected model.**

The p-values .....

Problem 5
---------

**Construct a 95% one step ahead forecast interval for the log exchange rate, based on your ARIMA-ARCH model.**

```{r}
coef(fit.var)
(omega <- coef(fit.var)[[1]])
(alpha <- coef(fit.var)[-1])

observations <- (resid^2)[1215:1219]

(h1 <- omega + sum(alpha * observations))

f1 <- as.numeric(forecast(fit.mean, h=1)$mean)

# ARIMA-ARCH inteval
f1 + sqrt(h1) * c(-1.96, 1.96) # 95%
```

**Compare this to the interval based on the ARIMA only model from Problem 2.**

Problem 6
---------

Here are the conditional variances, with the last 10 values printed out:

```{r}
ht <- fit.var$fit[,1]^2 # ARIMA-ARCH
tail(ht, n=10)
```

Here is a plot of the conditional variances:

```{r}
par(mfrow=c(1,2))
plot(date, diff.log.tsla, type="l", col=4)
plot(date, ht, type="l", col=4)
```

**Use this plot to locate bursts of high volatility.**

**Do these highly volatile periods agree with those found from examination of
the time series plot of the log exchange rates themselves?**

Problem 7
---------

Here is a time series plot which simultaneously shows the log exchange rates,
together with the ARIMA-GARCH one-step-ahead 95% forecast intervals based on
information available in the previous day:

```{r fig.width=12, fig.height=8}
par(mfrow=c(1,1))
plot(date, log.tsla, type="l")
lines(date, f + 1.96 * sqrt(ht), lty=2, col=2)
lines(date, f - 1.96 * sqrt(ht), lty=2, col=2)
```

**Using this plot, comment on the accuracy and practical usefulness of the
forecast intervals.**

Problem 8
---------

Here is a normal probability plot of the ARCH residuals.

```{r}
resid.arch <- resid / sqrt(ht)

qqnorm(resid.arch, col=2)
qqline(resid.arch, col=1, lty=2)
```

**Does the model seem to have adequately described the leptokurtosis ("long-tailedness") in the data?**

Problem 9
---------

Here is a count of how many prediction interval failures there were:

```{r}
sum(abs(log.tsla - f) > sqrt(ht) * 1.96, na.rm=TRUE)
```

The number of prediction intervals is:
```{r}
sum(!is.na(resid.arch))
60/1214
```

**What percentage of the time did the intervals fail?

About 5% of all intervals failed at forecasting the values.  




Sys.getenv("PATH")
#                                                      PATH 
#"/usr/bin:/bin:/usr/sbin:/sbin:/usr/local/bin:/usr/texbin" 
x <- matrix(1:6, nrow=2, dimnames=list(c('a','b'),c('c','d','this that')))
latex(x)






