Homework 6
==========
*Firstname Lastname (Replace this part with your name)*

```{r}
library("forecast")
```

For this assignment, we analyze the `chaos1` and `chaos2` series.

```{r}
chaos1 <- read.csv("http://ptrckprry.com/course/forecasting/data/chaos1.csv")$value
chaos2 <- read.csv("http://ptrckprry.com/course/forecasting/data/chaos2.csv")$value
time <- 1:50
```


Problem 1
---------

```{r}
x <- .500

f1 <- function(x) {
  x <- x/.6
  print(x)
}

f2 <- function(x) {
  x <- ((1-x)/.4)
  print(x)
}


for (i in 1:50) {
  if (0 <= x & x <= .6) 
    x <- f1(x)
  else 
    x <- f2(x)
  }


```

**Check that `x1 = f(x0)`, for `chaos1`.**

x1 matches f(x0).  The results are off by a small margin of error.  The last observation in chaos1 is 0.99152051 while my check has a final value of 0.98681.  The values digress later in the dataset.  


Problem 2
---------

```{r}
# Plot chaos1 and chaos2 in separate plots.
par(mfrow=c(1,1))
plot(time, chaos1, type = 'b', col = 4)

par(mfrow=c(1,1))
plot(time, chaos2, type = 'b', col = 4)
```

**Do the series look random?**

Chaos1 does not appear to be entirely random.  There seems to be some structure in the plot where the observation f1(x) drops close to zero and gradually works it way up to 1 before it drops again.  These drops in f1(x) occur at a constant cadence as well. 

Chaos2 appears to be more random versus Chaos1. The f2(x) output does not drop close zero at a consistant cadence like f1(x).

**Are they in fact random?**

No they are not random.  Both plots have time series patterns.  In both plots, there are tendencies for both plots to trend upward overtime, until it reaches a shock and drops drastically. Because of this pattern and structure, I conclude that these models are not random.  

http://web.ipac.caltech.edu/staff/fmasci/home/statistics_refs/TestForRandomness_RunsTest.pdf

**Do the series look stationary?**

The models do look stationary overtime.  

Problem 3
---------

```{r}
library("forecast")

par(mfrow=c(1,2))
Acf(chaos1)
Pacf(chaos1)

```

**Based on these, suggest an ARMA model.**

Based on this observation, I would suggest an ARIMA(2,0,0).  The ACF gradually dies down until (somewhat) but clearly cuts off in the PACF.  

**Would this model provide the best possible forecasts?**

??? 

Problem 4
---------

```{r}
par(mfrow=c(1,1))
plot(time, chaos1, 
     type = 'b', col = 2)
lines(time, chaos2, 
      type = 'b', col = 4)

legend("topleft", inset=0.05,
       legend=c("chaos1", "chaos2"), lty = c(1,1), col=c("red","blue"))
```

**Do the paths look similar?**

The plots look similar in terms of structure over time.  The difference is the lag in between shocks between both plots.  

**Should they look similar when t is close to 1?**

There should be a shock and they both f(x) values drop close to zero.  That should happen to both plots considering the parameters of f(x).  

**What should happen if `chaos1` and `chaos2` happen to get very close
together at some later time?**


Problem 5
---------

```{r}

xt1 <- chaos1[-50]
xt2 <- chaos1[-1]

plot(xt1, xt2, type = "p")

```


**Does this plot reveal the map (in other words, the function *f*) which
generated the data?**  

Yes this plot reveals the function.  


**Do you see why this *f* is called the tent map?**


**Does this plot help us to see that `{ x_t }` is not an AR(1) series?  How?**

