Homework 8
==========
*Kenneth Luna*
https://talksonmarkets.files.wordpress.com/2012/09/time-series-analysis-with-arima-e28093-arch013.pdf


For this assignment, we will use the `forecast` and `tseries` libraries.

```{r}
library("forecast")
library("tseries")
```

The dataset is the exchange rate for the Indian Rupee to 1 U.S. Dollar.
The data is daily from July 1st, 2002 to April 8th, 2011.

```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/rupee.csv")
date <- as.Date(data$date, format="%m/%d/%Y")
time <- 1:length(date)
rupee <- data$rupee
```

We will work with the logs of the exchange rates:

```{r}
log.rupee <- log(rupee)
```


Problem 1
---------

Here is a time series plot of log.rupee, along with the ACF and PACF:

```{r}
par(mfrow=c(1,1))
plot(time, log.rupee, type="l",
     xlab="Date", ylab="log.rupee")

par(mfrow=c(1,2))
Acf(log.rupee)
Pacf(log.rupee)
```

Here is the first difference, along with the ACF and PACF:

```{r}
## Add code to compute the first difference and make the plots.
diff.log.rupee <- c(NA, diff(log.rupee))

par(mfrow=c(1,1))
plot(time, diff.log.rupee, type="l",
     xlab="Date", ylab="1st-Difference Rupee")

par(mfrow=c(1,2))
Acf(diff.log.rupee)
Pacf(diff.log.rupee)

```

Here is the second difference, along with the ACF and PACF:

```{r}
## Add code to compute the second difference and make the plots.
diff2.log.rupee <- c(NA, diff(diff.log.rupee))

par(mfrow=c(1,1))
plot(time, diff2.log.rupee, type="l",
     xlab="Date", ylab="1st-Difference Rupee")

par(mfrow=c(1,2))
Acf(diff2.log.rupee)
Pacf(diff2.log.rupee)
```

**Does the series appear to be stationary?**

The series does not appear to be stationary.  The mean and the variance appears to change over time.  If we look at the plot for the first difference and second difference, the values tend to deviate from historical observations.  

**Can you identify an ARIMA(p,d,q) model from these plots?**

This model is an ARMA(1,0) model based on the ACF and PACF charts.

Problem 2
---------

Here are the AICc for all ARIMA(p,1,q) models without constant for p and q
ranging from 0 to 2:


```{r}

d <- 1
aiccMin <- 0
# choose p, q with AICc
for (include.constant in c(FALSE, TRUE)) {
    for (p in 0:4) {
        for (q in 0:4) {
            # work-around bug in R by manually differencing
            fit <- Arima(diff(log.rupee), c(p,0,q), include.constant=include.constant, method="ML")
            cat("ARIMA(", p, ",", d, ",", q, ")(constant=", include.constant, ") : ", fit$aicc, "\n", sep="");
    
            if (fit$aicc < aiccMin)
              aiccMin <- fit$aicc      
        }
    }
}


print(aiccMin)

# Add code to compute the AICc values.  You can modify the code from
# http://ptrckprry.com/course/forecasting/lecture/nasdaq-arch.html
# if you don't want to do this by hand.
```

**Select an ARIMA(p,1,q) model.**

The best fitting model is the ARIMA(1,1,0) based on the AICC value of -17763.37.  

Here is code to fit the model, then compute residuals and the fitted values:

```{r}
(fit.mean <- Arima(log.rupee, c(1,1,0), include.constant=FALSE, method="ML"))
```

Here are the residuals, with the last 10 residuals printed out:
```{r}
resid <- as.numeric(residuals(fit.mean))
tail(resid, n=10)
```

Here are the fitted values, with the last 10 fitted values printed out:
```{r}
f <- fitted.values(fit.mean)
tail(f, n=10)
```

Here is the one step ahead forecast and 95% forecast interval:

```{r}
# Add code to compute the forecast and interval.  You can do this
# by hand, or you can use the "forecast" function.
forecast(fit.mean, h = 1, level = 95)

```


Problem 3
---------

Here is a plot of the residuals:

```{r}
# Add plot of the residuals.
par(mfrow=c(1,1))
plot(time, resid, type="l")

```

Here are an ACF and PACF of the residuals:

```{r}
par(mfrow=c(1,2))
Acf(resid)
Pacf(resid)
```

Here are an ACF and PACF of the squared residuals:

```{r}
par(mfrow=c(1,2))
Acf(resid^2)
Pacf(resid^2)
```

**Use these plots to argue that the residuals, although approximately
uncorrelated, are not independent; instead, they show evidence of conditional
heteroscedasticity.**

These plots show a change in variance overtime, there indicates conditional heteroscedasticity.  Furthmore, the ACF and PACF for the resid and resid-squared shows that there continues to be signaificant amounts of variance at later lags.  


Problem 4
---------

Here are the AICc values for the ARCH(q):

```{r}
# Add code to compute the AICc values for the ARCH(q) models.
# You can modify the code from 
# http://ptrckprry.com/course/forecasting/lecture/nasdaq-arch.html
q <- 0:10
loglik <- rep(NA, length(q))
N <- length(resid)

for (i in 1:length(q)) {
    if (q[i] == 0) {
        loglik[i] <- -0.5 * N * (1 + log(2 * pi * mean(resid^2)))
    } else {
        fit <- garch(resid, c(0,q[i]), trace=FALSE)
        loglik[i] <- logLik(fit)
    }
}

k <- q + 1
aicc <- -2 * loglik  + 2 * k * N / (N - k - 1)

print(data.frame(q, loglik, aicc))

```

Here is the AICc for the GARCH(1,1):

```{r}
# Add code to compute the AICc values for the GARCH(1,1) models.
# You can modify the code from 
# http://ptrckprry.com/course/forecasting/lecture/nasdaq-arch.html
fit <- garch(resid, c(1,1), trace=FALSE)
loglik <- logLik(fit)
k <- 2
aicc <- -2 * loglik  + 2 * k * N / (N - k - 1)

print(data.frame(loglik, aicc))

```


Here are the summary and log likelihood of the selected model:

```{r}
fit.var <- garch(resid, c(1,1), trace=FALSE)
summary(fit.var)
logLik(fit.var)
```

**Comment on the statistical significance of the parameter values of your
selected model.**

The p-values for all of the coefficients is less than .05, therefore indicating that they are all statistically signifcant.  

**Write the complete form of the ARCH or GARCH model you have selected.**
http://en.wikipedia.org/wiki/Autoregressive_conditional_heteroskedasticity



Problem 5
---------

**Construct a 95% one step ahead forecast interval for the log exchange rate,
based on your ARIMA-ARCH model.**

```{r}
# You will need to to this by hand.  First, get the forecast (conditional
# mean) from the end of Problem 2.  Store this in a variable called "f1":
# f1 <- ?????
f1 <- as.numeric(forecast(fit.mean, h=1)$mean)

# Next, you need to determine the conditional variance, which we will store
# in a variable called "h1".
(omega <- coef(fit.var)[[1]])
(alpha <- coef(fit.var)[[2]])
(beta <- coef(fit.var)[[3]])
ht <- fit.var$fit[,1]^2 # ARIMA-ARCH
# To do this, use the output from Problem 4 to determine the coefficients
# (omega, alpha, etc.).  Use the output from Problem 2 to get the most
# recent residuals (shocks).  If your model is GARCH(1,1), you will also need
# the most recent conditional variance, which you can get from Problem 6.
# 
# Copy and paste the values and compute the conditional variance:
h1 <- omega + alpha * (resid^2)[2235] + beta * ht[2235]

# Finally, we compute the 95% forecast interval using the following command:
f1 + c(-1, 1) * 1.96 * sqrt(h1)
#forecast(fit.mean, h = 1, level = 95)
```

**Compare this to the interval based on the ARIMA only model from Problem 2.**

The ARIMA-GARCH forecast is more specific versus the ARIMA model.  The forecast intervals deviate with the data instead of expanding like the ARIMA model.  


Problem 6
---------

Here are the conditional variances, with the last 10 values printed out:

```{r}
ht <- fit.var$fit[,1]^2 # ARIMA-ARCH
tail(ht, n=10)
```

Here is a plot of the conditional variances:

```{r}
par(mfrow=c(1,2))
plot(date, diff.log.rupee, type="l", col=4)
plot(date, ht, type="l", col=3)
```

**Use this plot to locate bursts of high volatility.**


**Do these highly volatile periods agree with those found from examination of
the time series plot of the log exchange rates themselves?**


Problem 7
---------

Here is a time series plot which simultaneously shows the log exchange rates,
together with the ARIMA-ARCH one-step-ahead 95% forecast intervals based on
information available in the previous day:

```{r fig.width=12, fig.height=8}
par(mfrow=c(1,1))
plot(date, log.rupee, type="l")
lines(date, f + 1.96 * sqrt(ht), lty=2, col=2)
lines(date, f - 1.96 * sqrt(ht), lty=2, col=2)
```

**Using this plot, comment on the accuracy and practical usefulness of the
forecast intervals.**



Problem 8
---------

Here is a normal probability plot of the ARCH residuals.

```{r}
# Add code to compute the arch residuals:
resid.arch <- resid / sqrt(ht)

# Now, add code to make a normal probability plot (with the qqnorm command)
qqnorm(resid.arch, col=2)
qqline(resid.arch, col=1, lty=2)
```

**Does the model seem to have adequately described the leptokurtosis
("long-tailedness") in the data?**


Problem 9
---------

Here is a count of how many prediction interval failures there were:

```{r}
# The following command counts the number of arch residuals bigger than
# 0.5.  Modify this command to count the number of times the prediciton
# interval failed:
sum(abs(log.rupee - f) > sqrt(ht) * 1.96, na.rm=TRUE)
```

The number of prediction intervals is:
```{r}
# Uncomment to get a count of the number of non-missing arch residuals:
sum(!is.na(resid.arch))
```

**What percentage of the time did the intervals fail?

